---
title: "iteration_listcol"
author: "Caleigh Dwyer"
date: "2023-11-01"
output: github_document
---

basic structure of a for loop:

input = list (...)
output = list(...)

for (i in 1:n) {
output [[i]] = f(input[[i]])
}

However, for loops can get tedious and a little opaque. Loop functions are more efficient. They don't use indices. we'll use purrr::map(). Don't use Base R for loop functions in homework

basic structure of map:

output = map(input, f)

produces the same result as a for loop, and isn't any faster, but it is cleaner/clearer in your write up

workflow:
write a single example for fixed inputs, then embed example in a for loop, then abstract loop body to a function, then rewrite using a map statement

lists: help to store things that don't fit in data rectangles (can keep track of different types of inputs or outputs)

a dataframe is a special kind of list

A list can even contain a list! If you have an input list and output list of the same length, you can store that in a dataframe. benefit of keeping it in a dataframe is that you can do as many tidyverse transformations as you want.

you can even have a list storing your dataframes

```{r}
library(tidyverse)
library(rvest)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##Lists
You can put anything in a list

```{r}
l = list(
vec_numeric = 5:8,
vec_logical = c(TRUE, TRUE, FALSE, TRUE, FALSE, FALSE),
mat = matrix(1:8, nrow = 2, ncol= 4),
summary = summary(rnorm(100))
)
```


```{r}
l

l$vec_numeric

l[[1]]

l[["vec_numeric"]]

##all of these call the same things

mean(l[["vec_numeric"]])
```

##for loop

create a new list

```{r}
list_norm = 
  list(
    a= rnorm(20, mean=3, sd =1),
    b= rnorm(30, mean=0, sd =5),
    c= rnorm(40, mean=10, sd =.2),
    d= rnorm(20, mean=-3, sd =1)
  )
```


```{r}
list_norm
```

```{r}
mean_and_sd = function(x){
  
  if(!is.numeric(x)){
    stop("Input must be numeric")
  } 
  if (length(x)<3){
    stop("Input must have at least 3 numbers")
  }
  mean_x = mean(x)
  sd_x = sd(x)
  
  tibble(
    mean= mean_x,
    sd = sd_x
  )
  
}
```


```{r}
mean_and_sd(list_norm[[1]])
mean_and_sd([[2]])
mean_and_sd([[3]])
mean_and_sd([[4]])

##even thought you have a function, you have to keep applying it to each different input. you don't want to do that. so we're moving into for loop
```

let's use a for loop

```{r}
output = vector("list", length = 4)

output[[1]] = mean_and_sd(list_norm[[1]]) ##but we don't want to have to keep copying and pasting this

for (i in 1:4){
  output[[i]] = mean_and_sd(list_norm[[i]])
}

##this has now computed the mean and sd for each element of the list and pasted it into the output!
```


##let's try map

```{r}
output = map(list_norm, mean_and_sd)

##map just did the same thing as the for loop! but kept the vector name for each vector
```


what if you want a different function?

```{r}
output = map(list_norm, median)

output = map(list_norm, IQR)


```


```{r}
output = map_dbl(list_norm, median)
##the output is a little clearer here

output = map_df(list_norm, mean_and_sd)

output = map_df(list_norm, mean_and_sd, .id = "input")
##this adds an ID column to the df

```


##List columns!

```{r}
listcol_df =
  tibble(
    name = c("a", "b", "c", "d"),
    samp= list_norm
  )
```


```{r}
listcol_df %>% pull(name)
listcol_df %>% pull(samp)

listcol_df%>%
  filter(name=="a")
```


```{r}
listcol_df$samp[[1]]
mean_and_sd(listcol_df$samp[[1]])

```


```{r}
map(listcol_df$samp, mean_and_sd)

##not really sure what's going on with the listcol here
```

so can i add a list column?

```{r}
listcol_df =
listcol_df %>%
  mutate(summary = map(samp, mean_and_sd),
medians = map_dbl(samp, median))
```


##weather data

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728", "USW00022534", "USS0023B17S"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2021-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(
      id, 
      USW00094728 = "CentralPark_NY", 
      USW00022534 = "Molokai_HI",
      USS0023B17S = "Waterhole_WA"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```


```{r}
weather_nest = 
  weather_df |> 
  nest(data = date:tmin)
```


```{r}
weather_nest |> pull(name)
weather_nest |> pull(data)

weather_nest$data[[3]]
```


Suppoise i want to regress tmax on tmin for each station. 

```{r}
lm(tmax~tmin, data = weather_nest$data[[1]])
```


let's write a function.

```{r}
weather_lm = function(df){
  
  lm(tmax~tmin, data = df)
}

weather_lm(weather_nest$data[[1]])

output = vector("list", 3)

for(i in 1:3){
  output[[i]] = weather_lm(weather_nest$data[[i]])
}

##this allows you to write a regression for three dataframes at once


```

now as a map

```{r}
map(weather_nest$data, weather_lm)
```


```{r}
weather_nest=
weather_nest |> 
  mutate(models = map(data, weather_lm))

weather_nest$models
```

