---
title: "iteration_functions"
author: "Caleigh Dwyer"
date: "2023-10-26"
output: html_document
---

```{r}
library(tidyverse)
library(ggridges)
library(patchwork)
library(rvest)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


```{r}
set.seed(12345)

##this ensures that every time you get random numbers, it uses the same set of random numbers
```


## Z score function
z scores subtract the mean and divide by the sd

```{r}
x_vec = rnorm(20, mean = 5, sd = .3)
```

compute z scores for x_vec

```{r}
(x_vec - mean(x_vec))/sd(x_vec)

```

write this as a function to be applied to other vectors:

```{r}
z_score = function(x){
  
  z= (x-mean(x))/sd(x)
  
  z
}
```


```{r}
z_score(x=x_vec)

# this runs the function we just made

z_score(x = rnorm(10, mean = 5))
```


Keep checking

```{r}
z_score(x = 3)

z_score(c("my", "name", "is", "jeff"))

z_score(c(TRUE, TRUE, FALSE, TRUE))

z_score(iris)

##none of these work, for various reasons. you can't run it on characters
```

Change the function

```{r}
z_score = function(x){
  
  if(!is.numeric(x)){
    stop("Argument should be numbers")
  } else if (length(x)<2){
    stop("You need at least 2 numbers to get z scores")
  }
  z= (x-mean(x))/sd(x)
  
  z
}
```


let's try again:

```{r}
z_score(x = 3)

z_score(c("my", "name", "is", "jeff"))

z_score(c(TRUE, TRUE, FALSE, TRUE))

z_score(iris)

##none of these work, but now our output tells us why this doesn't work because we've used conditional statements that explicitly name the requirements
```


##multiple outputs

write a function that returns the mean and sd from a sample of numbers

```{r}
mean_and_sd = function(){
   
  if(!is.numeric(x)){
    stop("Argument should be numbers")
  } else if (length(x)<2){
    stop("You need at least 2 numbers to get z scores")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)
  
  tibble(
    mean = mean_x,
    sd = sd_x
  )
  
  
}
```


double check i did this right...

```{r}
mean_and_sd(x_vec)

#not working, says there's an unused argument
```



###start getting means and sds

```{r}

x_vec = rnorm(n = 30, mean = 5, sd = .5)

tibble(
  mean = mean(x_vec),
  sd = sd(x_vec)
)

```

let's write a function that uses "n" a true mean and a true SD as inputs

```{r}
sim_mean_sd = function(n_obs, mu, sigma) {
  
  x_vec = rnorm(n = n_obs, mean = mu, sd = sigma)

tibble(
  mean = mean(x_vec),
  sd = sd(x_vec)
)
}

sim_mean_sd(n_obs = 3000, mu = 50, sigma = 12.3)

##now you have one line of code that can produce many different outputs

sim_mean_sd(12, 24, 4)
```



first step is to check that your input argument output work:

```{r}
 fellowship_ring = readxl::read_excel("./data/LotR_Words.xlsx", range = "B3:D6") |>
  mutate(movie = "fellowship_ring")

##it does
```


write a function that at least has the cell range and movie title. will also put in path in case path gets renamed


```{r}

lotr_function = function(path = "data/LotR_Words.xlsx", cell_range, movie_name){
  movie_df = 
    readxl::read_excel(path, range = cell_range) |> 
    mutate(movie = movie_name)
  
  movie_df

}

##not working because of "attempt to use zero length variable name"? Fixed -- accidentally broke the chunk

lotr_function(cell_range = "B3:D6", movie_name = "fellowship_ring")
lotr_function(cell_range = "F3:H6", movie_name = "two_towers")
lotr_function(cell_range = "J3:L6", movie_name = "two_towers")
```

```{r}
lotr_function = function(path = "data/LotR_Words.xlsx", cell_range, movie_name){
  movie_df = 
    readxl::read_excel(path, range = cell_range) |> 
    janitor::clean_names() |> 
    gather(key = sex, value = words, female:male) |> 
             mutate(race = str_to_lower(race),
                    movie = movie_name)
  
  df

}

##used "cell_range" to name argument in function to differentiate from the argument "range" that is specific to the read_excel command
```


```{r}

you can write on multiple lines at once by using "shift + option" and drag to the number of rows you want to write on
you can write on multiple lines at once by using "shift + option" and drag to the number of rows you want to write on
you can write on multiple lines at once by using "shift + option" and drag to the number of rows you want to write on
```



```{r}
nsduh_url = "http://samhda.s3-us-gov-west-1.amazonaws.com/s3fs-public/field-uploads/2k15StateFiles/NSDUHsaeShortTermCHG2015.htm"

nsduh_html = read_html(nsduh_url)

data_marj = 
  nsduh_html |> 
  html_table() |> 
  nth(1) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent)) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
```

let's do this as a function instead

```{r}
nsduh_import = function(html, table_number, outcome_name){
  html |> 
  html_table() |> 
  nth(table_number) |>
  slice(-1) |> 
  select(-contains("P Value")) |>
  pivot_longer(
    -State,
    names_to = "age_year", 
    values_to = "percent") |>
  separate(age_year, into = c("age", "year"), sep = "\\(") |>
  mutate(
    year = str_replace(year, "\\)", ""),
    percent = str_replace(percent, "[a-c]$", ""),
    percent = as.numeric(percent),
    outcome = outcome_name) |>
  filter(!(State %in% c("Total U.S.", "Northeast", "Midwest", "South", "West")))
  
}

nsduh_import(html = nsduh_html, table_number = 1, outcome_name = "marj")
nsduh_import(html = nsduh_html, table_number = 4, outcome_name = "cocaine")
```


